### 2022-06-14 paper.tex
variably-methylated -> variably methylated

\begin{minipage}{0.7\textwidth}
\small
\textit{Abstract} -- Single-cell-bisulfite sequencing (scBS) allows for assessing DNA methylation with single-basepair and single-cell resolution. Analysing the big data sets thus obtained requires preprocessing to reduce data size, improve signal over noise, and provide interpretability. Typically, this is simply done by dividing the genome into large tiles and averaging in these. We show that this coarse-graining dilutes signal, and suggest improved approaches to determine more informative regions quantify methylation in, and a better quantitation method than simple averaging. We demonstrate how our approach enables better distinguishing of cell types and reduces demands on cell number. We also show that our method of finding variably-methylated regions finds biologically meaningful loci. Finally, we present a software tool, \texttt{scbs}, that implements this approach and offers further functionality required to work with scBS data.
\end{minipage}
\vspace{6.5ex}

@@

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.png}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \texttt{scbs} package.}
		\texttt{scbs prepare} parses methylation files produced by common bisulfite sequencing mappers and stores their contents in a compressed format optimised for efficient access to genomic intervals.
		\texttt{scbs} also produces cell-wise summary statistics and quality plots (here: average methylation around the transcription start site) that are used to detect low-quality cells.
		These cells can be discarded with \texttt{scbs filter}.
		To obtain a methylation matrix, similar to the count matrices used in scRNA-seq, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest, or they may choose to discover VMRs (variably-methylated regions) in the data with \texttt{scbs scan}.
		The resulting methylation matrix can then be used for downstream analysis such as cell clustering and dimensionality reduction.
	}
	\label{workflow}
\end{figure*}

@@

Here, we have proposed an improved strategy to preprocess single-cell bisulfite-sequencing data.
Based on the observation that incomplete read coverage of a genomic interval can lead to inaccurate methylation estimates, we suggest a scoring scheme that is aware of a reads local context rather than just treating all reads in an interval alike
Furthermore, we show a way to pinpoint minimal regions of high variability across cells, which we call variably-methylated regions (VMRs).
By reanalyzing two published datasets we could show that these improvements to data preprocessing help to increase signal and decrease noise, resulting in a more informative intermediate-dimensional representation of the data. As examples of practical benefits, we demonstrate that our preprocessing allows for better distinction of cell sub-types, especially for smaller data sets with limited numbers of cells. In addition to this improvement of global signal-to-noise ratio (with regards to cell dissimilarity quantification), we also show that the determination of VMRs improve biological interpretability by pinpointing the regions that actually change and hence have a putative regulatory role.

>>>>

\begin{minipage}{0.7\textwidth}
\small
\textit{Abstract} -- Single-cell-bisulfite sequencing (scBS) allows for assessing DNA methylation with single-basepair and single-cell resolution. Analysing the big data sets thus obtained requires preprocessing to reduce data size, improve signal over noise, and provide interpretability. Typically, this is simply done by dividing the genome into large tiles and averaging in these. We show that this coarse-graining dilutes signal, and suggest improved approaches to determine more informative regions quantify methylation in, and a better quantitation method than simple averaging. We demonstrate how our approach enables better distinguishing of cell types and reduces demands on cell number. We also show that our method of finding variably methylated regions finds biologically meaningful loci. Finally, we present a software tool, \texttt{scbs}, that implements this approach and offers further functionality required to work with scBS data.
\end{minipage}
\vspace{6.5ex}

@@

\begin{figure*}[t]
	\begin{center}
		\includegraphics[width=.8\textwidth]{figures/Fig_workflow.png}
	\end{center}
	\caption{\small \textbf{Overview of the functionalities implemented in the \texttt{scbs} package.}
		\texttt{scbs prepare} parses methylation files produced by common bisulfite sequencing mappers and stores their contents in a compressed format optimised for efficient access to genomic intervals.
		\texttt{scbs} also produces cell-wise summary statistics and quality plots (here: average methylation around the transcription start site) that are used to detect low-quality cells.
		These cells can be discarded with \texttt{scbs filter}.
		To obtain a methylation matrix, similar to the count matrices used in scRNA-seq, the user must first decide in which genomic intervals methylation should be quantified.
		The user may either provide genomic regions of \emph{a priori} interest, or they may choose to discover VMRs (variably methylated regions) in the data with \texttt{scbs scan}.
		The resulting methylation matrix can then be used for downstream analysis such as cell clustering and dimensionality reduction.
	}
	\label{workflow}
\end{figure*}

@@

Here, we have proposed an improved strategy to preprocess single-cell bisulfite-sequencing data.
Based on the observation that incomplete read coverage of a genomic interval can lead to inaccurate methylation estimates, we suggest a scoring scheme that is aware of a reads local context rather than just treating all reads in an interval alike
Furthermore, we show a way to pinpoint minimal regions of high variability across cells, which we call variably methylated regions (VMRs).
By reanalyzing two published datasets we could show that these improvements to data preprocessing help to increase signal and decrease noise, resulting in a more informative intermediate-dimensional representation of the data. As examples of practical benefits, we demonstrate that our preprocessing allows for better distinction of cell sub-types, especially for smaller data sets with limited numbers of cells. In addition to this improvement of global signal-to-noise ratio (with regards to cell dissimilarity quantification), we also show that the determination of VMRs improve biological interpretability by pinpointing the regions that actually change and hence have a putative regulatory role.
